\input{../../resources/lesson-head.tex}

\subsection*{Week 5 at a glance}

\subsubsection*{We will be learning and practicing to:}
%data types
%proof signposts
%using proofs to evaluate
%universal generalization
%applying proof strategy
%logical structure to proof strategy
%identifying proof strategy in proof
\begin{itemize}

\item Clearly and unambiguously communicate computational ideas using appropriate formalism. Translate across levels of abstraction.
\begin{itemize}
   \item Translating between symbolic and English versions of statements using precise mathematical language
    \item Using appropriate signpost words to improve readability of proofs, including 'arbitrary' and 'assume'
\end{itemize}

\item Know, select and apply appropriate computing knowledge and problem-solving techniques. Reason about computation and systems. Use mathematical techniques to solve problems. Determine appropriate conceptual tools to apply to new situations. Know when tools do not apply and try different approaches. Critically analyze and evaluate candidate solutions.
\begin{itemize}
    \item Judging logical equivalence of compound propositions using symbolic manipulation with known equivalences, including DeMorgan's Law
    \item Writing the converse, contrapositive, and inverse of a given conditional statement
    \item Determining what evidence is required to establish that a quantified statement is true or false
    \item Evaluating quantified statements about finite and infinite domains
\end{itemize}

\item Apply proof strategies, including direct proofs and proofs by contradiction, and determine whether a proposed argument is valid or not.
\begin{itemize}
    \item Identifying the proof strategies used in a given proof
    \item Identifying which proof strategies are applicable to prove a given compound proposition based on its logical structure
    \item Carrying out a given proof strategy to prove a given statement
    \item Carrying out a universal generalization argument to prove that a universal statement is true
    \item Using proofs as knowledge discovery tools to decide whether a statement is true or false
\end{itemize}
\end{itemize}

\subsubsection*{TODO:}
\begin{list}
   {\itemsep2pt}
   \item Project due May 7, 2024. No review quiz this week.
   \item Test 1, in class this week, on Friday May 3, 2024.
   The test covers material in Weeks 1 through 4 and Monday of Week 5. 
   To study for the exam, we recommend reviewing class notes 
   (e.g. annotations linked on the class website, podcast, supplementary video from the class website), 
   reviewing homework (and its posted sample solutions), and in particular *working examples* 
   (extra examples in lecture notes, review quizzes, discussion examples) and getting feedback (office hours and Piazza).
   Some practice questions (and their solutions) are available on the class website, linked from Week 5 and from the Assignments page.
\end{list}

\newpage

\begin{comment}
Removed definition of insertion, deletion, mutation from Wednesday of Week 4 -- when do we need them?
\input{../activity-snippets/algorithm-redundancy.tex}
\newpage
\input{../activity-snippets/cartesian-product-definition.tex}
\input{../activity-snippets/algorithm-rna-mutation-insertion-deletion.tex}
\input{../activity-snippets/rna-mutation-insertion-deletion-example.tex}
\end{comment}

\section*{Week 5 Monday: Nested Quantifiers}
\input{../activity-snippets/rna-rnalen-basecount-definitions.tex}
\input{../activity-snippets/alternating-quantifiers-order-rna-examples.tex}
\subsection*{Proof strategies}
\input{../activity-snippets/proof-strategies-quantification-finite-domain.tex}
%\input{../activity-snippets/proof-strategy-universal-exhaustion.tex}
\input{../activity-snippets/proof-strategy-universal-generalization.tex}
\newpage
\input{../activity-snippets/quiz-translating-counting-quantifiers.tex}

\newpage
\section*{Week 5 Wednesday: Proof Strategies and Sets}
\input{../activity-snippets/sets-equality-subset-definition.tex}
\input{../activity-snippets/proof-strategies-conditionals.tex}
\input{../activity-snippets/proof-strategies-proof-by-cases.tex}
\input{../activity-snippets/proof-strategies-ands.tex}
\input{../activity-snippets/sets-proof-strategies.tex}
\newpage
\input{../activity-snippets/sets-equality-example.tex}
\input{../activity-snippets/sets-basic-proofs.tex}
\vfill
\input{../activity-snippets/proofs-signposting.tex}
\newpage
\input{../activity-snippets/set-operations-union-intersection-powerset.tex}
\end{document}
