\input{../../resources/lesson-head.tex}

\subsection*{Week 8 at a glance}

\subsubsection*{We will be learning and practicing to:}
%classify cardinality
%important sets
%function and relation definitions
%functions for cardinality
%contradiction proofs
\begin{itemize}

\item Clearly and unambiguously communicate computational ideas using appropriate formalism. Translate across levels of abstraction.
\begin{itemize}
    \item Defining important sets of numbers, e.g. set of integers, set of rational numbers
    \item Defining functions using multiple representations
    \item Classifying sets into: finite sets, countably infinite sets, uncountable sets
    \item Using functions to compare cardinality of sets
\end{itemize}

\item Know, select and apply appropriate computing knowledge and problem-solving techniques. Reason about computation and systems. Use mathematical techniques to solve problems. Determine appropriate conceptual tools to apply to new situations. Know when tools do not apply and try different approaches. Critically analyze and evaluate candidate solutions.
\begin{itemize}
    \item Determining what evidence is required to establish that a quantified statement is true or false
    \item Evaluating quantified statements about finite and infinite domains
\end{itemize}

\item Apply proof strategies, including direct proofs and proofs by contradiction, and determine whether a proposed argument is valid or not.
\begin{itemize}
    \item Tracing and/or modifying a proof by contradiction
    \item Using proofs as knowledge discovery tools to decide whether a statement is true or false
\end{itemize}
\end{itemize}

\subsubsection*{TODO:}
\begin{list}
   {\itemsep2pt}
   \item Homework assignment 5 (due Tuesday May 21, 2024)
   \item Review quiz based on class material each day (due Friday May 24, 2024).
   \item Start reviewing for Test 2. The test is in class next week on Friday May 31, 2024.

\end{list}

\newpage

\section*{Week 8 Monday: Cardinality of Sets}

%\input{../activity-snippets/sets-numbers-subsets.tex}
\input{../activity-snippets/finite-sets-definition.tex}
\input{../activity-snippets/cardinality-motivation.tex}
\input{../activity-snippets/cardinality-rationale-for-functions.tex}
\input{../activity-snippets/musical-chairs-analogy.tex}
\newpage
\input{../activity-snippets/well-defined-functions.tex}
\newpage
\input{../activity-snippets/injective-function-definition.tex}
\input{../activity-snippets/injective-functions-visually.tex}
\input{../activity-snippets/cardinality-lower-bound-definition.tex}
\input{../activity-snippets/injective-cardinality-musical-chairs.tex}
%\input{../activity-snippets/rna-injective-cardinality.tex}
\newpage
\input{../activity-snippets/surjective-function-definition.tex}
\input{../activity-snippets/surjective-functions-visually.tex}
\input{../activity-snippets/cardinality-upper-bound-definition.tex}
\input{../activity-snippets/surjective-cardinality-musical-chairs.tex}
\input{../activity-snippets/bijection-definition.tex}

%\input{../activity-snippets/rna-surjective-cardinality.tex}

\newpage

\section*{Week 8 Wednesday and Friday: Finite, countably infinite, and uncountable sets}
\subsection*{Cardinality of sets}
\input{../activity-snippets/cardinality-definition.tex}
\input{../activity-snippets/cardinality-caution.tex}
\input{../activity-snippets/cardinality-properties.tex}
\input{../activity-snippets/cantor-schroder-bernstein-theorem.tex}
\newpage
\input{../activity-snippets/countably-infinite-definition.tex}
\input{../activity-snippets/countably-infinite-examples-sets-of-numbers.tex}
\input{../activity-snippets/countably-infinite-examples-other-sets.tex}
\newpage
\subsection*{Cardinality categories}
\input{../activity-snippets/cardinality-categories.tex}
\input{../activity-snippets/cardinality-countability-lemmas.tex}

\subsection*{Are there always *bigger* sets?}
\input{../activity-snippets/cardinality-power-sets.tex}
\newpage
\subsection*{Countable vs.\ uncountable: sets of numbers}
\input{../activity-snippets/cardinality-rationals-reals.tex}
\subsection*{Other examples of uncountable sets}
\input{../activity-snippets/cardinality-uncountable-examples.tex}
\newpage

\section*{Review Quiz}
\begin{enumerate}
    \item Sets of numbers. \hspace{1in}\\ \input{../activity-snippets/quiz-rationals-proofs.tex}
    \item Finite vs. infinite. \hspace{1in}\\ \input{../activity-snippets/quiz-finite-sets.tex}
    \item Functions. \hspace{1in}\\ \input{../activity-snippets/quiz-injective-surjective.tex}
    \item Functions. \hspace{1in}\\ \input{../activity-snippets/quiz-cardinality-witnessing-functions-q1.tex} 
    \item Functions. \hspace{1in}\\ \input{../activity-snippets/quiz-cardinality-witnessing-functions-q2.tex} 
    \item Diagonalization. \hspace{1in}\\ \input{../activity-snippets/quiz-diagonalization.tex}
    \item Classifying cardinality. \hspace{1in}\\ \input{../activity-snippets/quiz-cardinality-classifying.tex}
\end{enumerate}
\end{document}