\input{../../resources/lesson-head.tex}

\subsection*{Week 9 at a glance}

\subsubsection*{We will be learning and practicing to:}
%classify cardinality
%important sets
%function and relation definitions
%functions for cardinality
%div and mod
%divisibility and primes
%congruence mod n
%binary relations properties
%special binary relations
%graph representations of relations
\begin{itemize}

\item Clearly and unambiguously communicate computational ideas using appropriate formalism. Translate across levels of abstraction.
\begin{itemize}
    \item Defining important sets of numbers, e.g. set of integers, set of rational numbers
    \item Classifying sets into: finite sets, countably infinite sets, uncountable sets
    \item Defining functions, predicates, and binary relations using multiple representations
    \item Determining whether a given binary relation is symmetric, antisymmetric, reflexive, and/or transitive
    \item Determining whether a given binary relation is an equivalence relation and/or a partial order
\end{itemize}

\item Know, select and apply appropriate computing knowledge and problem-solving techniques. Reason about computation and systems. Use mathematical techniques to solve problems. Determine appropriate conceptual tools to apply to new situations. Know when tools do not apply and try different approaches. Critically analyze and evaluate candidate solutions.
\begin{itemize}
    \item Using the definitions of the div and mod operators on integers
    \item Using divisibility and primality predicates
    \item Applying the definition of congruence modulo n and modular arithmetic
\end{itemize}

\item Apply proof strategies, including direct proofs and proofs by contradiction, and determine whether a proposed argument is valid or not.
\begin{itemize}
    \item Using proofs as knowledge discovery tools to decide whether a statement is true or false
\end{itemize}
\end{itemize}

\subsubsection*{TODO:}
\begin{list}
   {\itemsep2pt}
   \item Review for Test 2. The test is in class on Friday May 31, 2024.
   \item Homework assignment 6 (due Thursday June 6, 2024).
\end{list}

\newpage
\section*{Week 9 Monday: No class in observance of Memorial Day}
\section*{Week 9 Wednesday: Binary relations}

\subsection*{Cardinality of sets: recap}
\input{../activity-snippets/cardinality-recap.tex}
\newpage
\subsection*{Binary relations}
\input{../activity-snippets/binary-relation-definition.tex}
\vfill
\input{../activity-snippets/relations-as-graphs.tex}
\vfill
\newpage
\input{../activity-snippets/binary-relation-examples.tex}
\newpage
\subsection*{Properties of binary relations}
\input{../activity-snippets/reflexive-relation-definition.tex}
\input{../activity-snippets/reflexive-relation-informally.tex}
\vfill
\input{../activity-snippets/symmetric-relation-definition.tex}
\input{../activity-snippets/symmetric-relation-informally.tex}
\vfill
\input{../activity-snippets/transitive-relation-definition.tex}
\input{../activity-snippets/transitive-relation-informally.tex}
\vfill
\input{../activity-snippets/antisymmetric-relation-definition.tex}
\input{../activity-snippets/antisymmetric-relation-informally.tex}
\vfill
\newpage
\input{../activity-snippets/binary-relation-properties-examples.tex}
\vfill

{\it Summary}: binary relations can be useful for organizing elements in a domain. 
Some binary relations have special properties that make them act like some familiar relations.
Equivalence relations (reflexive, symmetric, transitive binary relations) ``act like'' equals.
Partial orders (reflexive, antisymmetric, transitive binary relations) ``act like'' less than or equals to.


\end{document}